The Ecology of Cloud Computing

### 2022-04-07 paper.tex
Fixed grammar and expanded futurism and conclusion

On the other hand, we could host our website on the cloud by renting servers in a data center somewhere. For example, Amazon's Elastic Compute Cloud (EC2) charges \$0.10 per hour for each computer you require, and only for the hours where the computers are actively running. You can host your website off their datacenter, and Amazon has dedicated people who look after the servers and make sure they're running correctly. As you can see, this is a very attractive alternative to self-hosting, particularly at large scales. However, when you do this, it is easy to forget about the real, physical equipment behind whatever computing you're doing - without care, we can fall into seeing the cloud as an abstract \emph{thing} that processes our data for us, an amorphous blob of circuits beyond concrete concerns. Without mindfulness, the emissions problem can thus easily slip out of our hands.

\subsection{Some intuition for carbon use}
Most estimates of emissions are in the units CO2e, or "Carbon dioxide equivalent". This is to provide a standardised unit for emissions from different gases - for example, the climate impact of 1kg of methane is the same as the impact of 25kg of CO2 \citep{forster2007changes}, so if a process emits 3kg of methane and 4kg of CO2 it has emissions of $(3*25)+4=79$kg CO2e. Still, we don't generally have good intuition when it comes to knowing what 1g CO2e actually \emph{means}. Here are some benchmarks (the first three taken from \citet{lannelongue2021ten}):
\begin{itemize}
\item In a European car, driving one kilometre emits about 175g CO2e.
\item Flying from Paris to London emits around 50kg CO2e.
\item Streaming video for one hour emits around 55g CO2e (although this is hard to keep track of; electricity use from data transmission has halved every 2 years on average \citep{aslan2018electricity}).
\item The entire life-cycle of an iPhone 11 used for 3 years is 72kg CO2e (79\% of which is from manufacturing). \citep{apple2019product}
\item One kWh of electricity from the UK grid emits 233g CO2e. \citep{uk2020greenhouse}
\end{itemize}

@@

An NFT uses an algorithm to create non-fungible digital data - like a digital certificate of authenticity. It uses a technology called blockchain to make a digital receipt that cannot be copied or faked. A blockchain is essentially a list (of transactions) and a signature - the signature is a line of letters and numbers that encapsulates the contents of the list. If someone tampers with the list, or tries to improperly add something to it, the signature and list won't match and you will be able to tell it has been tampered with.\citep{narayanan2017bitcoin} This can be a good thing; artists have found use in it, being able to sell digital art as a limited-stock item (like traditional paintings are) directly, without than having to go through a middleman like an auction house. In fact, in 2020,  artists earned \$1.8 million \citep{redman2020NFT} from selling their art as NFTs.\newline

Now, where's the catch? The problem is how this uses blockchain. The generation of this signature takes a lot of computation; it is estimated that one NFT transaction generates about 48kg of CO2 \citep{atken2020unreasonable}, 14 times the footprint of mailing a physical art print, and 74000 times the footprint of buying something via bank transfer. \citep{qiu2021what} In fact, if you had turned on a laptop the day the Roman Empire was founded, and used it from then until December 2020, that laptop would've produced less carbon emissions than NFTs did in 2020. \citep{atken2020unreasonable} \newline

And all this is done for an algorithm that isn't even necessary; an alternate blockchain algorithm called proof-of-stake produces the exact same result with only 1\% of the computing power. \citep{saleh2021blockchain} This case study gives a stunning parable on how computing, what we use it for, and what algorithms we use can have a huge effect on the climate.

@@

With the ever-increasing processing power and memory of modern computers, it can be easy to not see efficiency as so mission-critical, especially for small algorithms - shaving one microsecond of runtime off a popular algorithm might not seem like much, but if that algorithm is used a trillion times (which isn't so far-fetched, when you consider algorithms like Fast Fourier Transforms or signal processing tools) you've saved 11.6 days of processing time (and 7kg CO2e) total.  As your program gets faster, there is also tendency to increase its scope and process larger data. This can create a 'rebound' effect \citep{lannelongue2021ten} that ends up turning your efficiency improvements into larger energy use. In this regard, restraint should be exercised, alongside careful consideration of how much data you actually \emph{need} to do good science.\newline

@@

This is probably the area where it is easiest to 'de-abstract' our thinking. As humans, we are used to thinking in space, and most of us in our daily lives run up to troubles with our phone, laptop or cloud service running out of space for us just as easily as we run out of space in a bookshelf or spare room. On a corporate scale, we may want to think about the actual usefulness of our archives, and 'declutter' as needed. Do we need to keep data from a 10-year-old study that has since far been superseded? Could we offload older data onto relevant parties (essentially, contacting researchers and saying 'we can give you your data if you like, but we aren't keeping it ourselves')? Maybe we could take a page from libraries and keep usage statistics, shedding datasets that haven't been touched or asked for in years? Perhaps in our heads we may do well to separate 'just in case!' from 'just for when...', and assess whether we are archiving or simply hoarding.

\subsection{Machine learning}
Machine learning is becoming \textit{en vogue} in scientific computing; it is being used for everything from improving early stage drug discovery \citep{imrie2021generating} to solving systems of differential equations. \citep{bhattacharya2020model} In essence, machine learning consists of two stages. The first is a 'training' stage, where a large dataset is processed by an algorithm in a machine; the second is a 'application' stage where the machine then uses what it has learned to make estimations about new data. For example, if I were training a machine to read handwritten letters (see \citet{deng2012mnist}), the 'training' dataset would be a big set of handwritten letters along with what they represent (so we'd tell the machine 'these 50 handwritten digits are all the letter 'a' etc.), and the machine would use an algorithm to figure out similarities and differences between letters. The 'application' stage would then be giving it a handwritten letter it hasn't seen before and it figuring out what letter it is from the data it gathered during the 'training' stage. This means machine learning can do a lot of busywork that we previously had to employ humans to do, like picking out good candidates for new drugs or typing up handwritten letters; essentially, it could be seen as a way of 'automating intuition'. \newline

This learning stage can take a lot of computation - the MNIST database from \citet{deng2012mnist} has 60,000 training letters. The Enron corpus (\citet[see][]{klimt2004enron}), a dataset of emails used for English language machine learning, contains 700,000 emails. And each of these needs an algorithm to be run on it individually. The total process can take days, and for advanced applications can use  huge amounts of electricity. \citet{strubell2019energy} found that training one model using Google's state-of-the-art BART machine learning framework takes 79 hours, in the process using 1507kWh\footnote{This calculation accounted for the cost of cooling and maintaining the data center during this time, not just the actual processing.} (approximately the energy used by a UK household in six months \citep{waters2019energy}) and 326kg CO2e (!). Smaller frameworks, like the T2T framework, used around 27kWh and 5kg CO2e (which is still a non-negligible amount). \newline

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we've generated over 5kg CO2e and \emph{we haven't even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job.

@@

This statistic can be a useful tool to consider priorities in where this simulation step can be leveraged. For example, if a department is trying to decide which simulation tool to put funding into, one with a 1:40 ratio may be more attractive than one with a 1:5 ratio. Although, one disadvantage is that this statistic doesn't mean anything on an objective level. For a molecular analysis experiment, 1:40 may be very attractive (as one could feasibly reap those time-saving benefits), whereas for a bioinformatics study the complexity of biological processes like protein interactions would make it a struggle to do a significant simulation in that time.

@@

\subsection{Data centres}
A key factor in considering cloud computing is the idea of 'overhead' - around 40\% of the energy use of a datacenter is purely the energy used to cool it so the servers don't overheat,  \citep{lannelongue2021ten} which would not be necessary in a local server (which is small enough to be cooled effectively by similar techniques as those used to cool human beings). However, as datacentres are geared purely towards the processing and storage of data, this can be done more efficiently. In fact, a 2013 case study (\citet{masanet2013energy}) suggested that by moving common applications used in a company from a local server to the cloud, they could reduce emissions by 87\%\footnote{Not to mention the further reduction in emissions if employees are then able to work from home; if it's all on the cloud anyway, then why would they need to burn petrol to commute?}. Naturally, one large cloud centre can utilise efficiency 'economies of scale' better than dozens of local, small ones. And over time, this has certainly shown; research by Google found that while data centre use increased by 550\% between 2010 and 2018, the amount of energy consumed by datacentres only increased by 6\%, meaning the proportion of global energy used by datacentres did not increase despite their much greater usage. \citep{masanet2020recalibrating}\newline

If using a commercial data centre, one can find a better datacentre by shopping around for better carbon options; with the choice of any cloud provider in the world, you also have the choice of what country's grid they're plugged into. For example, generating one kWh in Switzerland emits 12g CO2e, while in Australia it is 880g. \citep{grealey2021carbon}. Another statistic of data centre efficiency is its PUE (Power Usage Effectiveness), which measures efficiency of a datacentre via the following ratio:

@@

The PUE of a data centre can be reduced through various considerations; the first is how to manage heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, so on \citep{ospina2018cooling}) can be much more efficient. And then, design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

We can also integrate some interesting computing approaches to infrastructure to reduce PUE. \citet{evans2016deepmind} reduced the energy use for cooling at a Google datacentre by 40\% through training an AI on centre infrastructure data such as temperatures, power, pump speeds (for water cooling) and then allowing it to control the datacentre's cooling system.\newline

@@

And then there's the interplay of the datacentre areas vs the human areas; waste heat from servers can be used to heat up areas containing people rather than separate central heating. One incredibly simple way of doing this is to have areas designed for people on higher storeys, so the heat can rise and do as such. \newline

@@

\subsection{A note on carbon offsets}
It is tempting consider balancing out the extra electricity use of our datacentre with carbon offsetting. Carbon offsetting is the act of spending money on things with a negative carbon cost to counteract the emissions of a process. This ranges from practices like giving money to reforestation companies, to slightly more dubious things like investing in carbon sequestration research. \newline

Care must be taken when using carbon offsets. At best, they are an important part in a carbon-reduction portfolio, one which actually puts financial incentive on \emph{keeping} ecosystems rather than destroying them for profit. At worst, they are a way for an organisation to funnel money into reducing their climate guilt. It's important to deeply consider offsetting methods; for example, planting trees sounds like a brilliant idea (a mature tree sequesters 11kg of CO2 per year \citep{akbari2002shade}), but it takes 20 years before a tree is mature enough to start sequestering any significant amount of carbon \citep{ghussain2020biggest}. By this point, we may have long passed a point of no return for destructive climate change. And besides, are the partners we're using for offsetting reliable? Are they protecting those trees? If your trees catch fire or are eaten by a deer when they've just sprouted, your offset may go entirely unrealised. \newline

A great case study where badly thought-through tree planting has been destructive for the Earth was the UK Government's 1981 Forestry Grant Scheme. To increase the UK's timber supply, the UK Government offered huge tax concessions to people who planted trees. \citep{watkins1983public} This concession was incredibly attractive to the super-rich, and soon people started looking for large tracts of land to plant these trees. Eyes moved to the Flow Country in the Scottish Highlands - a vast peatland, the largest bog in Europe. Roughly 500,000 hectares (5000 square kilometres) were drained and planted with conifers before the ecological importance of the Flow Country was recognised and the scheme ended; this was eventually done for the protection of habitats. The other importance of the Flow Country, unknown at the time, was the incredible carbon storage potential of the peatland; peatland bogs only make up 3\% of the world's land area, but store up to a \emph{quarter} of all the world's soil carbon. \citep{turetsky2015global} And when they're drained, they emit carbon rather than storing it; roughly 2 billion \emph{tonnes} of carbon dioxide are emitted from drained peatlands, making 5\% of all global human-created greenhouse gas emissions; not to mention that the trees planted on top of it grew poorly in the soil, and were essentially useless as timber. A badly-planned and badly-informed piece of environmental policy in the 1980s thus left huge ecological damage that Scotland is still working on repairing even today. \citep{gewin2020peat} \newline

One could make the analogy that carbon offsetting is the climate equivalent of the payday loan. Sure, your carbon output will balance out (over infinite time), but if you can help it you'd do better to just leave the fossil fuels in the ground in the first place - and worse, a badly thought-out payday loan could put you in worse financial straits than you were in to start with.

@@

Another area of interest is quantum computing - this uses concepts from quantum physics to redo computing in a more efficient way 'from scratch'. The difference that this will make to the environment is completely unimaginable; it is estimated to be capable of quickly working through algorithms that normal computers wouldn't be able to do between now and the end of the universe, which completely changes the lens through which we see computing efficiency. However, we cannot tell if the infrastructure required will balance it out in early days; as it stands, quantum computers must be supercooled to work, which uses a massive amount of cooling energy even on a very small scale.

\section{Conclusion}
The main idea flowing underneath this report is one of care; one of mindfulness. There are countless ways to improve the efficiency of our computing and cloud facilities so long as one simply cares about them - the climate crisis is a modern, global elephant in the room and it should equally be an elephant in project planning. Somebody who does not care about the climate (at their own peril) could create scientific software, could hire servers in a datacentre, could perhaps even build their own datacentre, not considering these factors. But you, the reader, are more careful than that; we should be acutely aware of our electricity-based carbon emissions in the same way we are aware of the plastic we are putting in our general waste or fumes coming out of our cars. It shouldn't be an 'out of sight, out of mind' problem, but we know it is an inevitability that we will produce this waste as part of our lives. Thus, it shouldn't be all or nothing; we should make a careful assessment of where we can reduce our energy use, and where we cannot. We should let it sway our judgements and play a factor in our science and scientific software. That way, it is possible to use the technology of the 'information age' in a way that is beneficial for both people and the planet.

>>>>

On the other hand, we could host our website on the cloud by renting servers in a data center somewhere. For example, Amazon's Elastic Compute Cloud (EC2) charges \$0.10 per hour for each computer you require, and only for the hours where the computers are actively running. You can host your website off their datacenter, and Amazon has dedicated people who look after the servers and make sure they are running correctly. As you can see, this is a very attractive alternative to self-hosting, particularly at large scales. However, when you do this, it is easy to forget about the real, physical equipment behind whatever computing you are doing - without care, we can fall into seeing the cloud as an abstract \emph{thing} that processes our data for us, an amorphous blob of circuits beyond concrete concerns. Without mindfulness, the emissions problem can thus easily slip out of our hands.

\subsection{Some intuition for carbon use}
Most estimates of emissions are in the units CO2e, or "Carbon dioxide equivalent". This is to provide a standardised unit for emissions from different gases - for example, the climate impact of 1kg of methane is the same as the impact of 25kg of CO2 \citep{forster2007changes}, so if a process emits 3kg of methane and 4kg of CO2 it has emissions of $(3*25)+4=79$kg CO2e. Still, we generally do not have good intuition when it comes to knowing what 1g CO2e actually \emph{means}. Here are some benchmarks (the first three taken from \citet{lannelongue2021ten}):
\begin{itemize}
\item In a European car, driving one kilometre emits about 175g CO2e.
\item Flying from Paris to London emits around 50kg CO2e.
\item Streaming video for one hour emits around 55g CO2e (although this is hard to keep track of; electricity use from data transmission has halved every 2 years on average \citep{aslan2018electricity}).
\item The entire life-cycle of an iPhone 11 used for 3 years is 72kg CO2e (79\% of which is from manufacturing). \citep{apple2019product}
\item One kWh of electricity from the UK grid emits 233g CO2e. \citep{uk2020greenhouse}
\end{itemize}

@@

An NFT uses an algorithm to create non-fungible digital data - like a digital certificate of authenticity. It uses a technology called blockchain to make a digital receipt that cannot be copied or faked. A blockchain is essentially a list (of transactions) and a signature - the signature is a line of letters and numbers that encapsulates the contents of the list. If someone tampers with the list, or tries to improperly add something to it, the signature and list will not match and you will be able to tell it has been tampered with.\citep{narayanan2017bitcoin} This can be a good thing; artists have found use in it, being able to sell digital art as a limited-stock item (like traditional paintings are) directly, without than having to go through a middleman like an auction house. In fact, in 2020,  artists earned \$1.8 million \citep{redman2020NFT} from selling their art as NFTs.\newline

Now, where's the catch? The problem is how this uses blockchain. The generation of this signature takes a lot of computation; it is estimated that one NFT transaction generates about 48kg of CO2 \citep{atken2020unreasonable}, 14 times the footprint of mailing a physical art print, and 74000 times the footprint of buying something via bank transfer. \citep{qiu2021what} In fact, if you had turned on a laptop the day the Roman Empire was founded, and used it from then until December 2020, that laptop would have produced less carbon emissions than NFTs did in 2020. \citep{atken2020unreasonable} \newline

Finally, all this is done for an algorithm that may not even be necessary; an alternate blockchain algorithm called proof-of-stake produces the exact same result with only 1\% of the computing power. \citep{saleh2021blockchain} This case study gives a stunning parable on how computing, what we use it for, and what algorithms we use can have a huge effect on the climate.

@@

With the ever-increasing processing power and memory of modern computers, it can be easy to not see efficiency as so mission-critical, especially for small algorithms - shaving one microsecond of runtime off a popular algorithm might not seem like much, but if that algorithm is used a trillion times (which is not so far-fetched, when you consider algorithms like Fast Fourier Transforms or signal processing tools) you save 11.6 days of processing time (and 7kg CO2e) total.  As your program gets faster, there is also tendency to increase its scope and process larger data. This can create a 'rebound' effect \citep{lannelongue2021ten} that ends up turning your efficiency improvements into larger energy use. In this regard, restraint should be exercised, alongside careful consideration of how much data you actually \emph{need} to do good science.\newline

@@

This is probably the area where it is easiest to 'de-abstract' our thinking. As humans, we are used to thinking in space, and most of us in our daily lives run into trouble with our phone, laptop or cloud service running out of space for us just as easily as we run out of space in a bookshelf or spare room. On a corporate scale, we may want to think about the actual usefulness of our archives, and 'declutter' as needed. Do we need to keep data from a 10-year-old study that has since far been superseded? Could we offload older data onto relevant parties (essentially, contacting researchers and saying 'we can give you your data if you like, but we aren't keeping it ourselves')? Maybe we could take a page from libraries and keep usage statistics, shedding datasets that have not been touched or asked for in years? Perhaps in our heads we may do well to separate 'just in case!' from 'just for when...', and assess whether we are archiving or simply hoarding.

\subsection{Machine learning}
Machine learning is becoming \textit{en vogue} in scientific computing; it is being used for everything from improving early stage drug discovery \citep{imrie2021generating} to solving systems of differential equations. \citep{bhattacharya2020model} In essence, machine learning consists of two stages. The first is a 'training' stage, where a large dataset is processed by an algorithm in a machine; the second is a 'application' stage where the machine then uses what it has learned to make estimations about new data. For example, if I were training a machine to read handwritten letters (see \citet{deng2012mnist}), the 'training' dataset would be a big set of handwritten letters along with what they represent (telling the machine 'these 50 handwritten digits are all the letter 'a' etc.), and the machine would use an algorithm to figure out similarities and differences between letters. The 'application' stage would then be giving it a handwritten letter it hasn't seen before and it figuring out what letter it is from the data it gathered during the 'training' stage. This means machine learning can do a lot of busywork that we previously had to employ humans to do, like picking out good candidates for new drugs or typing up handwritten letters; essentially, it could be seen as a way of 'automating intuition'. \newline

This learning stage can take a lot of computation - the MNIST database from \citet{deng2012mnist} has 60,000 training letters. The Enron corpus (\citet[see][]{klimt2004enron}), a dataset of emails used for English language machine learning, contains 700,000 emails; each of these needs an algorithm to be run on it individually. The total process can take days, and for advanced applications can use  huge amounts of electricity. \citet{strubell2019energy} found that training one model using Google's state-of-the-art BART machine learning framework takes 79 hours, in the process using 1507kWh\footnote{This calculation accounted for the cost of cooling and maintaining the data center during this time, not just the actual processing.} (approximately the energy used by a UK household in six months \citep{waters2019energy}) and 326kg CO2e (!). Smaller frameworks, like the T2T framework, used around 27kWh and 5kg CO2e (which is still a non-negligible amount). \newline

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we generated over 5kg CO2e and \emph{not even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job.

@@

This statistic can be a useful tool to consider priorities in where this simulation step can be leveraged. For example, if a department is trying to decide which simulation tool to put funding into, one with a 1:40 ratio may be more attractive than one with a 1:5 ratio. Although, one disadvantage is that this statistic means nothing concrete on an objective level. For a molecular analysis experiment, 1:40 may be very attractive (as one could feasibly reap those time-saving benefits), whereas for a bioinformatics study the complexity of biological processes like protein interactions would make it a struggle to do a significant simulation in that time. \newline

It is important to remember that a lot of scientific equipment (such as the ISIS facility) will likely never be green until the entire national grid is green; they simply use too much power. We can try to reduce it where we can (via methods like this), but to disparage the electricity use of these facilities in general belies the huge environmental benefits created through the scientific studies they make possible; the research into things like superconductors \citep{shang2018time} or creating sustainable synthetic versions of natural resources that is only possible through huge facilities like this.

@@

\subsection{Data centres}
A key factor in considering cloud computing is the idea of 'overhead' - around 40\% of the energy use of a datacenter is purely the energy used to cool it so the servers do not overheat,  \citep{lannelongue2021ten} which would not be necessary in a local server (which is small enough to be cooled effectively by similar techniques as those used to cool human beings). However, as datacentres are geared purely towards the processing and storage of data, this can be done more efficiently. In fact, a 2013 case study (\citet{masanet2013energy}) suggested that by moving common applications used in a company from a local server to the cloud, they could reduce emissions by 87\%\footnote{Not to mention the further reduction in emissions if employees are then able to work from home; if everything is on the cloud anyway, then why would they need to burn petrol to commute?}. Naturally, one large cloud centre can utilise efficiency 'economies of scale' better than dozens of local, small ones. Over time, this has certainly shown; research by Google found that while data centre use increased by 550\% between 2010 and 2018, the amount of energy consumed by datacentres only increased by 6\%, meaning the proportion of global energy used by datacentres did not increase despite their much greater usage. \citep{masanet2020recalibrating}\newline

If using a commercial data centre, one can find a better datacentre by shopping around for better carbon options; with the choice of any cloud provider in the world, you also have the choice of what country's grid they are plugged into. For example, generating one kWh in Switzerland emits 12g CO2e, while in Australia it is 880g. \citep{grealey2021carbon}. Another statistic of data centre efficiency is its PUE (Power Usage Effectiveness), which measures efficiency of a datacentre via the following ratio:

@@

The PUE of a data centre can be reduced through various considerations; the first is how to manage heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, so on \citep{ospina2018cooling}) can be much more efficient. Design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

We can also integrate some interesting computing approaches to infrastructure to reduce PUE. \citet{evans2016deepmind} reduced the energy use for cooling at a Google datacentre by 40\% through training an AI on centre infrastructure data such as temperatures, power, pump speeds (for water cooling) and then allowing it to control the datacentre's cooling system. However, for 'smaller wins', we must consider the power use of our computing approaches; the actual energy use of training and running this AI may be non-negligible (although in this case, probably less than 40\% of the cooling bill) \newline

@@

Additionally, we can leverage the interplay of the datacentre areas vs the human areas; waste heat from servers can be used to heat up areas containing people rather than separate central heating. One incredibly simple way of doing this is to have areas designed for people on higher storeys, so the heat can rise and do as such. \newline

@@

\subsection{A note on carbon offsets}
It is tempting to consider balancing out the extra electricity use of our datacentre with carbon offsetting. Carbon offsetting is the act of spending money on things with a negative carbon cost to counteract the emissions of a process. This ranges from practices like giving money to reforestation companies, to slightly more dubious things like investing in carbon sequestration research. \newline

Care must be taken when using carbon offsets. At best, they are an important part in a carbon-reduction portfolio, one which actually puts financial incentive on \emph{keeping} ecosystems rather than destroying them for profit. At worst, they are a way for an organisation to funnel money into reducing their climate guilt. It is important to deeply consider offsetting methods; for example, planting trees sounds like a brilliant idea (a mature tree sequesters 11kg of CO2 per year \citep{akbari2002shade}), but it takes 20 years before a tree is mature enough to start sequestering any significant amount of carbon \citep{ghussain2020biggest}. By this point, we may have long passed a point of no return for destructive climate change. Besides, are the partners we use for offsetting reliable? Are they protecting those trees? If your trees catch fire or are eaten by a deer as saplings, your offset may go entirely unrealised. \newline

A great case study where badly thought-through tree planting has been destructive for the Earth was the UK Government's 1981 Forestry Grant Scheme. To increase the UK's timber supply, the UK Government offered huge tax concessions to people who planted trees. \citep{watkins1983public} This concession was incredibly attractive to the super-rich, and soon people started looking for large tracts of land to plant these trees. Eyes moved to the Flow Country in the Scottish Highlands - a vast peatland, the largest bog in Europe. Roughly 500,000 hectares (5000 square kilometres) were drained and planted with conifers before the ecological importance of the Flow Country was recognised and the scheme ended; this was eventually done for the protection of habitats.

The other importance of the Flow Country, unknown at the time, was the incredible carbon storage potential of the peatland; peatland bogs only make up 3\% of the world's land area, but store up to a \emph{quarter} of all the world's soil carbon. \citep{turetsky2015global} Worse still, when drained, they emit carbon rather than storing it; roughly 2 billion \emph{tonnes} of carbon dioxide are emitted from drained peatlands, making 5\% of all global human-created greenhouse gas emissions; not to mention that the trees planted on top of it grew poorly in the soil, and were essentially useless as timber. A badly-planned and badly-informed piece of environmental policy in the 1980s thus left huge ecological damage that Scotland is still working on repairing even today. \citep{gewin2020peat} \newline

One could make the analogy that carbon offsetting is the climate equivalent of the payday loan. Sure, your carbon output will balance out (over infinite time), but if you can help it it would be better to just leave the fossil fuels in the ground in the first place - and worse, a badly thought-out payday loan could put you in worse financial straits than you were in to start with.

@@

Another area of interest is quantum computing - this uses concepts from quantum physics to redo computing in a more efficient way 'from scratch'. Quantum computing involves the use of 'qubits', which take advantage of the uncertainty present in quantum-mechanical information to have more states than the 0 or 1 of a classical computing bit, and leverage this for incredibly fast computation. \citep{national2019quantum} The difference that this will make to the environment is completely unimaginable; it is estimated to be capable of quickly working through algorithms that normal computers could not do between now and the end of the universe, which completely changes the lens through which we see computing efficiency. \citep{gyongyosi2019survey} However, we cannot tell if the infrastructure required will balance it out in early days; as it stands, quantum computers must be supercooled to work, which uses a massive amount of cooling energy even on a very small scale. In fact, a sub-field of research into quantum computing is whether it is possible to make large quantum computers without the excessive heat causing the quantum components to fail. \citep{ikonen2017energy}\newline

One direct environmental application of quantum computing is its ability to solve complex problems in renewable energy. Controlling and balancing a far more decentralised national grid (i.e. one with thousands of distributed renewable energy sources, rather than a few dozen fossil fuel plants) is an incredibly complex problem that current supercomputers do not have the resources to solve; it is simply too many variables to keep track of to forecast future supply and demand like the grid currently does. Other energy problems include the modelling of battery designs; simulation (which is heavily based in keeping track of the positions and movements of thousands and thousands of molecules) is an area where quantum computing is likely to create huge speed impacts. \citep{giani2021quantum}

\section{Conclusion}
The main idea flowing through this report is one of care; one of mindfulness. There are countless ways to improve the efficiency of our computing and cloud facilities so long as one simply cares about them - the climate crisis is a modern, global elephant in the room and it should equally be an elephant in project planning. Somebody who does not care about the climate (at their own peril) could create scientific software, could hire servers in a datacentre, could perhaps even build their own datacentre, not considering these factors. But you, the reader, are more careful than that; we should be acutely aware of our electricity-based carbon emissions in the same way we are aware of the plastic we are putting in our general waste or fumes coming out of our cars. It shouldn't be an 'out of sight, out of mind' problem, but we know it is an inevitability that we will produce this waste as part of our lives.\newline

There is a default set of questions that we ask when we carry out a project. We ask 'is this going to be useful?', 'is this user-friendly?', or 'is my new building accessible to disabled people?'. Perhaps it is time to add a new question; "does my project work for the environment, or against it?" Can we at least work in harmony with environmental goals, or even better, contribute to them? You don't even have to care about the environment to do this - more eco-friendly businesses have better job satisfaction with less burnout and turnover \citep{yu2020hotels} \citep{reverte2016influence}, better perception from consumers \citep{miles1993eco}, and have an easier time recruiting new and better talent \citep{aranganathan2018green} - the latter two growing more and more important, as younger, more environmentally-conscious individuals begin to join the workforce (and consumer perception is, naturally, potential employee perception).\newline

Thus, it shouldn't be all or nothing; we should make a careful assessment of where we can reduce our energy use, and where we cannot. We should let it sway our judgements and play a factor in our science and scientific software. That way, it is possible to use the technology of the 'information age' in a way that is beneficial for both people and the planet.
